---
title: "mocapr"
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- badges: start -->
[![Travis build status](https://travis-ci.com/steenharsted/mocapr.svg?token=TWbxTsB8WvZpmvPSsk8j&branch=master)](https://travis-ci.com/steenharsted/mocapr)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/steenharsted/mocapr?branch=master&svg=true)](https://ci.appveyor.com/project/steenharsted/mocapr)
<!-- badges: end -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of `mocapr` is to provide R functions to import, plot, animate, and analyse motion capture data. The package is in **the very early stages of development** and is only **minimally effective** in the sense that it, at current, only supports import from [the Captury](http://thecaptury.com/) system. My intent is to make this package work with motion capture data from other sources as well, and I expect to add a function that imports data from [the Kinetisense](https://kinetisense.com/) system shortly. 

If you have motion capture data from other systems than the Captury, it should be possible to wrangle the data into a format that will allow you to use the functions in this package, as long as the data contains frame by frame *joint center positions*. If you are willing to share some sample data, I will be happy to make an attempt at writing an import function and include both the function and the sample data in this package. 

`mocapr` uses a series of tidyverse packages to import ([`readr`](https://github.com/tidyverse/readr), [`tidyr`](https://github.com/tidyverse/tidyr), [`dplyr`](https://github.com/tidyverse/dplyr), [`stringr`](https://github.com/tidyverse/stringr), [`forcats`](https://github.com/tidyverse/forcats)), plot ([`ggplot2`](https://github.com/tidyverse/ggplot2)), animate ([`gganimate`](https://github.com/thomasp85/gganimate)), and analyse motion capture data.  
The package also contains a sample data set `mocapr_data` which is generated using some of the above packages as well as [`purrr`](https://github.com/tidyverse/purrr).

While all functions should run without loading other libraries I strongly recommend you load the tidyverse `library(tidyverse)` prior to loading the mocapr library.

This is my first R package, and I'd be surprised if have gotten everything right in the first try. Feedback and suggestions for improvement and future developments are **most welcome**. 

## Short on the story behind the package
My Ph.D. project is a part of the [MiPS](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5576290/) project, where we have followed a cohort of ~950 pre-school children with yearly follow-ups. Each follow-up consisted of a series of tests and examinations with one them being a series of squats and jumps recorded using markerless motion capture. My Ph.D. project revolves around collecting and analyzing this motion capture data, and the code available in this package is a more general form of some of the code I have written so far in my project.
My initial attempts at approaching the data were made in Stata, but my work significantly picked up speed after I was introduced to R in the summer of 2017, and especially after I discovered the Tidyverse.


## Installation
`mocapr` can  be installed directly from github using devtools:

```{r installation, eval=FALSE}
# install.packages('devtools')
devtools::install_github('steenharsted/mocapr')
```


## The Sample Data `mocapr_data`
`mocapr_data` consists of `r length(unique(mocapr::mocapr_data$movement_nr))` movements, each supplied with a number (`movement_nr`) and a short description (`movement_description`). All data included in the package is, at current, captured using the Captury Live markerless motion capture system. Raw exports in .csv format can be found in the folder "data-raw". Videos of the movements with an overlay of the track is available at this [YouTube playlist Link](https://www.youtube.com/playlist?list=PLMjrjny4Ymmd1nSGHU0A6dWfEWjBxc-VQ)

```{r load_packages_show_data}
suppressPackageStartupMessages(library(tidyverse))
library(mocapr)

#Data
mocapr::mocapr_data %>% 
  group_by(movement_nr, movement_description) %>% 
  tidyr::nest()
```

Each movement contains frame by frame joint angles and global joint center positions. All joint related variables are abbreviated according to their side (L|R), joint(A|K|H|S|E|W), and angle/position. As such each joint is typically represented by 6 columns.   

| Side    | Joint        |Angle/Position |
| :---    | :---         | :---          |    
|         | A  (Ankle)   |F  (Flexion)   |
|L (left) | K  (Knee)    |Varus          |
|         | H  (Hip)     |DF (Dorsi Flexion)|
|R (Right)| W  (Wrist)   |X  (joint center position on the global X axis (floor)|
|         | E  (Elbow)   |Y  (joint center position on the global Y axis)(up)|
|         | S  (Shoulder)|Z  (joint center position on the global Z axis)(floor)|


Example for left knee:

| Abbreviated Variable Name | Meaning of abbreviation |
| :---:        |     :---:      | 
|LKF | Left Knee Flexion | 
|LKX | Left Knee joint center position on the X axis (floor plane) | 


The focus of this tutorial is on plotting and animating motion capture data. For this we only need the joint center positions. I will not discuss the joint angles further, but feel free to explore them on your own.

Lets first create some sample data:
```{r create_data}
jump_1 <- mocapr::mocapr_data %>% 
  filter(movement_nr == 1)

jump_2 <- mocapr::mocapr_data %>% 
  filter(movement_nr == 2)

gait <-  mocapr::mocapr_data %>% 
  filter(movement_nr == 4)

caipoera <- mocapr::mocapr_data %>% 
  filter(movement_nr == 6)
```


The global joint center positions can be used for plots and animations using the `animate_global()` function.
```{r jump_1_GP}
jump_1 %>% 
  animate_global(nframes = nrow(.), fps = 50)
```
Plots and animations using global joint center positions will cause a oblique viewpoint if the subject moves at an angle to the axis of the global coordinate system. For the purpose of analyzing or interpreting motions an oblique viewpoint is, in general, less optimal.  
When I performed `jump_2` I simulated a poor landing on the right knee and made sure that the movement was oblique to global coordinate system. 
```{r jump_2_GP}
jump_2 %>% 
  animate_global(nframes = nrow(.), fps = 50)
```

In many cases out of axis movement is easy to prevent, and here the `animate_global()` function might be sufficient, but in other cases (e.g. working with pre-school children or more complicated movements than walking), out of axis movement is difficult to prevent - at least without interfering in the spontaneous movements of the subject. Also the the orientation of the global coordinate system differs from system to system and sometimes even between different setups of the same motion-capture system. This creates a need for animation and plotting functions that are free from the orientation of the global coordinate system, and instead focused on the subject or the direction of the movement the subject is performing.
`mocapr` solves this challenge by providing two functions that projects the global joint center positions onto the planes of the movement direction (`project_full_body_to_MP()`) or the anatomical planes the subject (`project_full_body_to_AP()`). Each project function should then be followed by it's corresponding animation function (`animate_movement()` or `animate_anatomical()`)

The movement planes are:  
* Forward) a plane perpendicular to the floor, going in the direction from the position of the subject at the first frame to the position of the subject at the last frame.  
* Side-wards) a plane perpendicular to the floor and the forwards plane.  

The anatomical planes are:  
* Frontal) a plane perpendicular to the floor, going through both hip-joint centers.  
* Sagital) a plane perpendicular to the floor and the Frontal plane.  

For movements where the subject is moving in one direction without rotation (such as walking in a straight line, or jumping using both legs) these two projections will be very similar to each other, but they will differ greatly if the direction of the movement changes.This is best explained by watching the animations produced by the animation functions `mocapr::animate_movement()` and `mocapr::animate_anatomical()`.

Lets look again at jump_2 (the jump that is oblique to the global coordinate system), and animate the jump using the `animate_movement()` and the `animate_anatomical()` functions.
```{r jump_2_MP}
jump_2 %>%
  #Project to the movements planes
  project_full_body_to_MP() %>%
  #Animate the movement plane projections
  animate_movement(nframes = nrow(.), fps = 50, rewind = FALSE)
```

```{r jump_2_AP}
jump_2 %>% 
  #Project to the anatomical Planes 
  project_full_body_to_AP() %>% 
  #Animate the anatomical projections
  animate_anatomical(nframes = nrow(.), fps = 50, rewind = FALSE)
```

Besides the size difference the two animations are very similar. This is because the movement that the subject is performing is uni-directional.   
*note: the right side appears on the right side in the anatomical animation and on the left side in the movement animation, this is intentional but might change in future versions*.  
Lets explore the difference between the two types of projections by looking at a movement that is not unidirectional.   
```{r walking_square_MP}
gait %>%
  #Project to the movements planes
  project_full_body_to_MP() %>%
  #Animate the movement plane projections
  animate_movement(nframes = nrow(.), fps = 50, height = 800, width = 800)
```

```{r walking_square_AP}
gait %>% 
  #Project to the anatomical Planes 
  project_full_body_to_AP() %>% 
  #Animate the anatomical projections
  animate_anatomical(nframes = nrow(.), fps = 50, height = 800, width = 800)
```

Now the difference between the two types of animations is evident. While both the animate_movement() and the animate_anatomical() gives you two view points that are perpendicular to each other, animate_movement() gives you *fixed viewpoints* (you are standing still and watching the movement) and animate_anatomical() *updates your viewpoint for each frame* (you are always the watching the subject from the front and the side of the pelvis).


## using `mocapr` to plot
The three animate functions can be used to plot if you supply the argument `animate = FALSE`. In that case the you will get a plot that is faceted on the frames. I suggest you reduce the number of frames before you use the functions to plot.

```{r}
jump_2 %>% 
  project_full_body_to_AP() %>% 
  filter( frame == 120 | frame == 150 | frame == 165 | frame == 170) %>% 
  animate_anatomical(animate = FALSE)
```

