---
title: "mocapr"
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`mocapr` uses a series of tidyverse packages to import ([`readr`](https://github.com/tidyverse/readr), [`tidyr`](https://github.com/tidyverse/tidyr), [`dplyr`](https://github.com/tidyverse/dplyr), [`stringr`](https://github.com/tidyverse/stringr), [`forcats`](https://github.com/tidyverse/forcats)), plot ([`ggplot2`](https://github.com/tidyverse/ggplot2)), animate ([`gganimate`](https://github.com/thomasp85/gganimate)), and analyse motion capture data.  
The package also contains sample data set `MOCAP_data` which is generated using some of the above packages as well as [`purrr`](https://github.com/tidyverse/purrr).

While all functions should run without loading other libraries I strongly recommend you load the tidyverse `library(tidyverse)`.


The current state of the package is **work in progress**.  
Feedback and suggestions for improvements are most welcome. 

## Why this package?
The field of motion capture has expanded rapidly over the last years
Motion capture data has become readily availble, and more will come
It is becoming easy to collect this type of data
Most systems are using their own export formats (especially true for markerless motion capture), and 
Open source tools to work with this data has not followed the same trajectory.

## The Vision

## Can you help?
Yes. Clinical background, ended up with the programming part by nessescity. No programming skills when I started the Ph.D., and I consider myself a beginner/intermediate R-user. While I think the code is correct, I am quite sure that it is not optimal. Input....

Data + video(if possible)
Function to import data into the `mocapr`format
global coordinate system
anatomical projections
movement projections



## Installation
`mocapr` can  be installed directly from github using devtools:

```{r installation, eval=FALSE}
# install.packages('devtools')
devtools::install_github('steenharsted/mocapr')
```


##The Data
All data currently obtained from the CapturyLive system
Kinetisense to come
Vicon plug-in-gait likely to come



```{r}
library(mocapr)
library(dplyr)
#Data
mocapr::MOCAP_data %>% 
  group_by(movement_nr, movement_description) %>% 
  tidyr::nest()

Jump <- mocapr::MOCAP_data %>%
  filter(movement_nr == 1)
```

##Example 1 
```{r}
#Project to movements planes
Jump %>% 
  project_full_body_to_MP() %>%
  #Animate the movement plane projections
  animate_movement(nframes = nrow(.)*2, fps = 50, rewind = FALSE)
```

##Example 2
```{r}
#Project to Anatomical Planes 
Jump %>% 
  project_full_body_to_AP() %>% 
  #Animate the anatomical projections
  animate_anatomical(nframes = nrow(.)*2, fps = 50, rewind = FALSE)
```

```{r}
library(dplyr)
Jump %>% 
  project_full_body_to_AP() %>% 
  filter(frame == 1 | frame == 50 | frame == 75| frame == 100 | frame == 150 | frame == max(frame)) %>% 
  animate_anatomical(animate = FALSE)
```

```{r}
library(dplyr)
Jump %>% 
  project_full_body_to_MP() %>% 
  filter(frame == min(frame) | frame == max(frame) | frame == 50) %>% 
  animate_movement(animate = FALSE)
```



