---
title: "mocapr"
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`mocapr` uses a series of tidyverse packages to import ([`tidyr`](https://github.com/tidyverse/tidyr), [`dplyr`](https://github.com/tidyverse/dplyr), [`stringr`](https://github.com/tidyverse/stringr), [`forcats`](https://github.com/tidyverse/forcats)), plot ([`ggplot2`](https://github.com/tidyverse/ggplot2)), animate ([`gganimate`](https://github.com/thomasp85/gganimate)), and analyse motion capture data.  

The current state of the package is **work in progress**.

##Why this package?
Motion capture data has become readily availble, and more will come
It is becoming easy to collect this type of data
Most systems are using their own export formats (especially true for markerless motion capture), and 

##The Vision

##Can you help?
Yes. Clinical background, ended up with the programming part by nessescity. No programming skills when I started the Ph.D., and I consider myself a beginner/intermediate R-user. While I think the code is correct, I am quite sure that it is not optimal. Input....

Data + video(if possible)
Function to import data into the `mocapr`format


## Installation
`mocapr` can  be installed directly from github using devtools:

```{r installation, eval=FALSE}
# install.packages('devtools')
devtools::install_github('steenharsted/mocapr')
```


##The Data
All data currently obtained from the CapturyLive system
Kinetisense to come
Vicon plug-in-gait likely to come

```{r}
library(mocapr)

#Data
Jump <- mocapr::jump_1
head(Jump)
```



##Example 1
```{r}
#Project to Anatomical Planes 
Jump <- cbind(Jump,
              #Upper extremity
              Project_to_AP(Jump, Y=LSY, X=LSX, Z=LSZ, New_Name = "LS"),
              Project_to_AP(Jump, Y=LEY, X=LEX, Z=LEZ, New_Name = "LE"),
              Project_to_AP(Jump, Y=LWY, X=LWX, Z=LWZ, New_Name = "LW"),
              Project_to_AP(Jump, Y=RSY, X=RSX, Z=RSZ, New_Name = "RS"),
              Project_to_AP(Jump, Y=REY, X=REX, Z=REZ, New_Name = "RE"),
              Project_to_AP(Jump, Y=RWY, X=RWX, Z=RWZ, New_Name = "RW"),
              #Lower extremity
              Project_to_AP(Jump,Y=LKY, X=LKX, Z=LKZ, New_Name = "LK"),
              Project_to_AP(Jump,Y=LHY, X=LHX, Z=LHZ, New_Name = "LH"),
              Project_to_AP(Jump,Y=LAY, X=LAX, Z=LAZ, New_Name = "LA"),
              Project_to_AP(Jump,Y=LTY, X=LTX, Z=LTZ, New_Name = "LT"),
              Project_to_AP(Jump,Y=RKY, X=RKX, Z=RKZ, New_Name = "RK"),
              Project_to_AP(Jump,Y=RHY, X=RHX, Z=RHZ, New_Name = "RH"),
              Project_to_AP(Jump,Y=RAY, X=RAX, Z=RAZ, New_Name = "RA"),
              Project_to_AP(Jump,Y=RTY, X=RTX, Z=RTZ, New_Name = "RT"))

#Animate the anatomical projections
Jump %>% animate_anatomical(nframes = nrow(.)*2, fps = 50, rewind = FALSE)
```

##Example 2
```{r}
#Project to movements planes
Jump <- cbind(Jump,
              #Upper extremity
              Project_to_MP(Jump, Y=LSY, X=LSX, Z=LSZ, New_Name = "LS"),
              Project_to_MP(Jump, Y=LEY, X=LEX, Z=LEZ, New_Name = "LE"),
              Project_to_MP(Jump, Y=LWY, X=LWX, Z=LWZ, New_Name = "LW"),
              Project_to_MP(Jump, Y=RSY, X=RSX, Z=RSZ, New_Name = "RS"),
              Project_to_MP(Jump, Y=REY, X=REX, Z=REZ, New_Name = "RE"),
              Project_to_MP(Jump, Y=RWY, X=RWX, Z=RWZ, New_Name = "RW"),
              #Lower extremity
              Project_to_MP(Jump,Y=LKY, X=LKX, Z=LKZ, New_Name = "LK"),
              Project_to_MP(Jump,Y=LHY, X=LHX, Z=LHZ, New_Name = "LH"),
              Project_to_MP(Jump,Y=LAY, X=LAX, Z=LAZ, New_Name = "LA"),
              Project_to_MP(Jump,Y=LTY, X=LTX, Z=LTZ, New_Name = "LT"),
              Project_to_MP(Jump,Y=RKY, X=RKX, Z=RKZ, New_Name = "RK"),
              Project_to_MP(Jump,Y=RHY, X=RHX, Z=RHZ, New_Name = "RH"),
              Project_to_MP(Jump,Y=RAY, X=RAX, Z=RAZ, New_Name = "RA"),
              Project_to_MP(Jump,Y=RTY, X=RTX, Z=RTZ, New_Name = "RT"))

#Animate the movement plane projections
Jump %>% animate_movement(nframes = nrow(.)*2, fps = 50, rewind = FALSE)
```



